---
title: "alaska gaby"
author: "Gaby Garcia"
date: "11/1/2019"
output: pdf_document
---

#Obtain Alaskan Data
```{r}
library(dplyr)
library(dataRetrieval)
AlaskaSites<-whatNWISsites(stateCd="02",parameterCd="00060")
```


## Bin width
```{r}
(max(AlaskaSites$dec_lat_va)-min(AlaskaSites$dec_lat_va))/10
```

## Assign Bin Width Value to a Variable
```{r}
AlaskaBin<-2.002339
```

### Check levels and change site_tp_cd to a factor
```{r}
levels(AlaskaSites$site_tp_cd)
AlaskaSites$site_tp_cd<-as.factor(AlaskaSites$site_tp_cd)
```

### Check class of site_tp_cd
```{r}
class(AlaskaSites$site_tp_cd)

```


### Filter AlaskaSites by Stream
```{r}
AlaskaSites<-filter(AlaskaSites, site_tp_cd=="ST")
```

### Read in Alaska Streamflow Sites document 
```{r, warning=FALSE, message=FALSE}
setwd("/Users/gabrielagarcia/Desktop/Hydrologic Data Analysis/Hydrologic_Data_Analysis/Course_Project/Alaska/DATA/PROCESSED")
Alaska_Sites <- read.csv("/Users/gabrielagarcia/Desktop/Hydrologic Data Analysis/Hydrologic_Data_Analysis/Course_Project/Alaska/DATA/PROCESSED/Alaska_Streamflow_Sites.csv")
```


## Filter Alaska Streamflow Sites document for Latitude Bin
```{r}
AlaskaLatitude5 <- filter(Alaska_Sites, Latitude_Bin == 5)
```
>Treat Latitude and longitude as continuous variable: how is repsonse variable changing with change in bin? Look at slope. 

```{r}
library(dataRetrieval)
date_summary <- AlaskaLatitude5$site_no
size <- vector(length = 1)
counter = 1
for(site in AlaskaLatitude5$site_no){
  flowdata <- readNWISdv(site, "00060")
  size[counter] <- length(flowdata$Date)
  counter = counter + 1
}
sitesize <- data.frame(cbind(AlaskaLatitude5$site_no, size))
sitesize$V1[which.max(sitesize$size)]
```
>Site with longest period of record in Bin 5 is 15276000, or SHIP C NR ANCHORAGE AK. Period of Record is 33 years. 


## Site Discharge Information 
```{r}
library(dataRetrieval)
AnchorageDischarge<- readNWISdv(siteNumbers = "15276000",
                     parameterCd = "00060",
                     startDate = "",
                     endDate = "")
```
>	readNWISdv returns time-series data summarized to a day. Default is mean daily.

### Write Site Discharge to a CSV
```{r}
Discharge<-write.csv(AnchorageDischarge, file = "Bin5Discharge.csv",row.names=FALSE)
```


## Site Information Summary
```{r}
 AnchorageSummary <- whatNWISdata(siteNumbers = "15276000")
```
>whatNWISdata is the most general NWIS data import function. User must explicitly define the service parameter. More flexible than the other functions.

### Write Site Information to a CSV
```{r}
AnchorageInformation<-write.csv(AnchorageSummary, file = "Bin5Information.csv",row.names=FALSE)
```


### Site Metadata
```{r}
AnchorageCodesCoordinates<-readNWISsite(siteNumber="15276000")
```
>Gives us drainage area, HUC Code, latitude and longitude 






































## Upload Alaskan Bin 5 Sites with periods of record included
```{r}
library(openxlsx)
AlaskaBin5<-read.xlsx("/Users/gabrielagarcia/Desktop/Hydrologic Data Analysis/Hydrologic_Data_Analysis/Course_Project/Alaska/DATA/RAW/USGSDischargeSitesBin5Final.xlsx")
```

### Find Period of Record
```{r}
library(dplyr)
AlaskaBin5 <- mutate(AlaskaBin5, PeriodofRecord = PeriodofRecordEnd-PeriodofRecordStart)
```

### Change Period of Record to factor
```{r}
AlaskaBin5$PeriodofRecord<-as.factor(AlaskaBin5$PeriodofRecord)
```

### Check levels of Period of Record
```{r}
levels(AlaskaBin5$PeriodofRecord)
```

>Longest period of Record for Bin 5 is CHILKAT R NR KLUKWAN AK, or 15056500



# Import all Data Frames
```{r, warning=FALSE}
setwd("/Users/gabrielagarcia/Desktop/Hydrologic Data Analysis/Hydrologic_Data_Analysis/Course_Project/Alaska/DATA/RAW")
NOAATempPrecipBin1<-read.csv("NOAA_Temp_Bin1.csv")
NOAATempPrecipBin2<-read.csv("NOAA_Temp_Bin2.csv")
NOAATempPrecipBin3<-read.csv("NOAA_Temp_Bin3.csv")
NOAATempPrecipBin5<-read.csv("Bin5TempandPrecip.csv")

NOAATempPrecip48910<-read.csv("NOAA_Precip_Temp_Bin_4_8_9_10.csv")
```

### Add Bin 2 Column to NOAATempPrecipBin2
```{r}
NOAATempPrecipBin2$Bin<-2
```


### Remove Tavg Column from Precip 
```{r, warning=FALSE, message=FALSE}
library(dplyr)
NOAATempPrecip48910<-select(NOAATempPrecip48910,-c(8))
```

### Add Bin Columns with values for 4, 8, and 9
```{r}
library(plyr)
NOAATempPrecip48910$Bin<-revalue(NOAATempPrecip48910$STATION,
                              c('USW00026510'='4', 'USC00500464'='8','USW00027502'='9'))
```

## Upload Bins 6 and 7 
```{r, message=FALSE, warning=FALSE}
setwd("/Users/gabrielagarcia/Desktop/Hydrologic Data Analysis/Hydrologic_Data_Analysis/Course_Project/Alaska/DATA/PROCESSED")
NOAATempPrecipBin67<-read.csv("NOAA_Temp_Precip_Bin_6_7.csv")
```

### Remove Snow Column from Precip Temp 6 & 7
```{r}
NOAATempPrecipBin67<-select(NOAATempPrecipBin67,-c(8))
```


## Fix Date Issues in NOAA Bin 1
```{r}
##Tell R to change factor to a date, in the format Month/day/year
NOAATempPrecipBin1$DATE <- as.Date(NOAATempPrecipBin1$DATE, format = "%m/%d/%y") 

# Changes the format of the "datetime" column to a two digit year, a two digit month, and the day as a number
NOAATempPrecipBin1$DATE <- format(NOAATempPrecipBin1$DATE, "%y%m%d")

#Creating a function that says that if d is greater than 181231, then it will return a 19, if d is not greater than 181231, then it will return a 20.Paste0 takes the two statements that are inside the function (d>181231, "19", "20")

#The 181231 is the last date of 2018. If d is over this value, a 19 is returned, If D<the number, a 20 is returned. 
create.early.dates <- (function(d) {
       paste0(ifelse(d > 181231,"19","20"),d)
       })


#This will input the datetime column into the function we created above for each row of the column
NOAATempPrecipBin1$DATE<- create.early.dates(NOAATempPrecipBin1$DATE)

#Changes the date time column into a different format to a 4 digit year, a two digit month, and the day as
NOAATempPrecipBin1$DATE <- as.Date(NOAATempPrecipBin1$DATE, format = "%Y%m%d") 

```

## Fix Date Issues in NOAA Bin 5
```{r}
##Tell R to change factor to a date, in the format Month/day/year
NOAATempPrecipBin5$DATE <- as.Date(NOAATempPrecipBin5$DATE, format = "%m/%d/%y") 

# Changes the format of the "datetime" column to a two digit year, a two digit month, and the day as a number
NOAATempPrecipBin5$DATE<- format(NOAATempPrecipBin5$DATE, "%y%m%d")

#Creating a function that says that if d is greater than 181231, then it will return a 19, if d is not greater than 181231, then it will return a 20.Paste0 takes the two statements that are inside the function (d>181231, "19", "20")

#The 181231 is the last date of 2018. If d is over this value, a 19 is returned, If D<the number, a 20 is returned. 
create.early.dates <- (function(d) {
       paste0(ifelse(d > 181231,"19","20"),d)
       })


#This will input the datetime column into the function we created above for each row of the column
NOAATempPrecipBin5$DATE<- create.early.dates(NOAATempPrecipBin5$DATE)

#Changes the date time column into a different format to a 4 digit year, a two digit month, and the day as
NOAATempPrecipBin5$DATE <- as.Date(NOAATempPrecipBin5$DATE, format = "%Y%m%d") 
```


## RBind all TempPrecip Data
```{r}
TempPrecip<-rbind(NOAATempPrecipBin1, NOAATempPrecipBin2, NOAATempPrecipBin3, NOAATempPrecipBin5, NOAATempPrecipBin67, NOAATempPrecip48910)
```


# Import in USGS Discharge Data
```{r, warning=FALSE, message=FALSE}
setwd("/Users/gabrielagarcia/Desktop/Hydrologic Data Analysis/Hydrologic_Data_Analysis/Course_Project/Alaska/DATA/RAW")
NWIS_Discharge<-read.csv("NWIS_Discharge.csv")
```

###Convert site_no to a factor
```{r}
NWIS_Discharge$site_no<-as.factor(NWIS_Discharge$site_no)
```

## Add Bins to Discharge Data
```{r}
library(plyr)
NWIS_Discharge$Bin<-revalue(NWIS_Discharge$site_no,
                              c('15297680'='1', '15297610'='2','15072000'='3', '15052500'='4', '15276000'='5', '15304000'='6', '15514000'='7', '15453500'='8', '15908000'='9', '15896000'='10'))
```

##Rename "Date" to "DATE" in NWIS_Discharge dataframe so both columns match
```{r}
names(NWIS_Discharge)[3]<-"DATE"
```


# Join TempPrecip and NWIS_Discharge Data Frames by Date and Bin 
```{r}
AlaskaTempPrecipDischarge <- TempPrecip %>%
  merge(NWIS_Discharge, by = c('DATE', 'Bin')) %>%
  rename(Discharge = X_00060_00003) %>%
  select(DATE:Discharge)
```



# Data Analysis and Visualization 

#  Precip


### Create Precip only Dataframe
```{r, message=FALSE, warning=FALSE}
library(dplyr)
AlaskaPrecipOnly<-select(AlaskaTempPrecipDischarge, .data$DATE, .data$PRCP, .data$Bin)
```

###Check number of Precipitation NA's and remove them
```{r}
sum(is.na(AlaskaPrecipOnly))
newdata<-na.omit(AlaskaPrecipOnly)
```

# Subset the Data by Bin
```{r}
Bin1Precip<- newdata%>%filter(Bin==1)
Bin2Precip<-newdata%>%filter(Bin==2)
Bin3Precip<-newdata%>%filter(Bin==3)
Bin4Precip<-newdata%>%filter(Bin==4)
Bin5Precip<-newdata%>%filter(Bin==5)
Bin6Precip<-newdata%>%filter(Bin==6)
Bin7Precip<-newdata%>%filter(Bin==7)
Bin8Precip<-newdata%>%filter(Bin==8)
Bin9Precip<-newdata%>%filter(Bin==9)
=======
AlaskaTempPrecipDischarge<-write.csv(AlaskaTempPrecipDischarge, "AlaskaTempPrecipDischarge.csv", row.names = F)
>>>>>>> 1cc38df18fbe03fc1dd2154172895ebf3a4066a7
```



```{r}
##precip30<- newdata%>%filter(DATE >= as.Date('1995-06-04') & DATE <= as.Date('2019-11-02'))
```


### Set Personal Theme
```{r}
gabytheme <- theme_bw(base_size = 22) + 
  theme(plot.title=element_text(face="bold", size="26", color="IndianRed3", hjust=0.5),
        axis.title=element_text(size=22, color="black"),
axis.text = element_text(face="bold", size=16, color = "black"), 
panel.background=element_rect(fill="white", color="darkblue"), 
panel.border = element_rect(color = "black", size = 2),
legend.position = "top", legend.background = element_rect(fill="white", color="black"),
            legend.key = element_rect(fill="transparent", color="NA"))

theme_set(gabytheme)
```



# Plot Bin 1 Precip
```{r, fig.width=9, fig.height=3}
library(scales)
library(ggplot2)
Bin1PrecipPlot<- 
  ggplot(Bin1Precip, aes(x = DATE, y = PRCP)) +
  geom_line() +
  ggtitle("Alaska Precipitation over Time") +
  labs(x = "", y = expression("Precipitation (in"^3*"/s)"))+
  scale_x_date(labels=date_format("%Y"),
               breaks=date_breaks("1 year"))+
  gabytheme
print(Bin1PrecipPlot)

```

# Plot Bin 2 Precip over Time
```{r, fig.width=12}
library(scales)
library(ggplot2)
Bin2PrecipPlot<- 
  ggplot(Bin2Precip, aes(x = DATE, y = PRCP)) +
  geom_line() +
  ggtitle("Alaska Precipitation over Time") +
  labs(x = "", y = expression("Precipitation (in"^3*"/s)"))+
  scale_x_date(labels=date_format("%y"),
               breaks=date_breaks("1 year"))+
  gabytheme
print(Bin2PrecipPlot)
```
>There is a gap in precipitation data collection for Bin 2 from 1986-12-31 to 1995-10-01

# Plot Bin 3 Precip over Time
```{r, fig.width=17}
library(scales)
library(ggplot2)
Bin3PrecipPlot<- 
  ggplot(Bin3Precip, aes(x = DATE, y = PRCP)) +
  geom_line() +
  ggtitle("Alaska Precipitation over Time") +
  labs(x = "", y = expression("Precipitation (in"^3*"/s)"))+
  scale_x_date(labels=date_format("%y"),
               breaks=date_breaks("1 year"))+
  gabytheme
print(Bin3PrecipPlot)

```
>Much higher range in precip values-->more rain and snowfall?

# Plot Bin 4 Precip over Time
```{r, fig.width=17}
library(scales)
library(ggplot2)
Bin4PrecipPlot<- 
  ggplot(Bin4Precip, aes(x = DATE, y = PRCP)) +
  geom_line() +
  ggtitle("Alaska Precipitation over Time") +
  labs(x = "", y = expression("Precipitation (in"^3*"/s)"))+
  scale_x_date(labels=date_format("%y"),
               breaks=date_breaks("1 year"))+
  gabytheme
print(Bin4PrecipPlot)

```

# Plot Bin 5 Precip over Time
```{r, fig.width=17}
library(scales)
library(ggplot2)
Bin5PrecipPlot<- 
  ggplot(Bin5Precip, aes(x = DATE, y = PRCP)) +
  geom_line() +
  ggtitle("Alaska Precipitation over Time") +
  labs(x = "", y = expression("Precipitation (in"^3*"/s)"))+
  scale_x_date(labels=date_format("%y"),
               breaks=date_breaks("1 year"))+
  gabytheme
print(Bin5PrecipPlot)

```

# Plot Bin 6 Precip over Time
```{r, fig.width=17}
library(scales)
library(ggplot2)
Bin6PrecipPlot<- 
  ggplot(Bin6Precip, aes(x = DATE, y = PRCP)) +
  geom_line() +
  ggtitle("Alaska Precipitation over Time") +
  labs(x = "", y = expression("Precipitation (in"^3*"/s)"))+
  scale_x_date(labels=date_format("%y"),
               breaks=date_breaks("1 year"))+
  gabytheme
print(Bin6PrecipPlot)

```
# Plot Bin 7 Precip over time

```{r, fig.width=17}
library(scales)
library(ggplot2)
Bin7PrecipPlot<- 
  ggplot(Bin7Precip, aes(x = DATE, y = PRCP)) +
  geom_line() +
  ggtitle("Alaska Precipitation over Time") +
  labs(x = "", y = expression("Precipitation (in"^3*"/s)"))+
  scale_x_date(labels=date_format("%y"),
               breaks=date_breaks("1 year"))+
  gabytheme
print(Bin7PrecipPlot)
```


# Plot Bin 8 Precip over time

```{r, fig.width=17}
library(scales)
library(ggplot2)
Bin8PrecipPlot<- 
  ggplot(Bin8Precip, aes(x = DATE, y = PRCP)) +
  geom_line() +
  ggtitle("Alaska Precipitation over Time") +
  labs(x = "", y = expression("Precipitation (in"^3*"/s)"))+
  scale_x_date(labels=date_format("%y"),
               breaks=date_breaks("1 year"))+
  gabytheme
print(Bin8PrecipPlot)
```
# Plot Bin 9 Precip over time
```{r, fig.width=17}
library(scales)
library(ggplot2)
Bin9PrecipPlot<- 
  ggplot(Bin9Precip, aes(x = DATE, y = PRCP)) +
  geom_line() +
  ggtitle("Alaska Precipitation over Time") +
  labs(x = "", y = expression("Precipitation (in"^3*"/s)"))+
  scale_x_date(labels=date_format("%y"),
               breaks=date_breaks("1 year"))+
  gabytheme
print(Bin9PrecipPlot)
```
>A lot less precip at highest latitude 