---
title: "Precipitation_Analysis"
author: "Gaby Garcia"
date: "11/11/2019"
output: pdf_document
---
# Load in relevant packages and complete Cleaned Alaska Dataset
```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
library(scales)
library(trend)
setwd("/Users/gabrielagarcia/Desktop/Hydrologic Data Analysis/Hydrologic_Data_Analysis/Course_Project/Alaska/DATA/PROCESSED")
AlaskaTempPrecipDischarge<-read.csv("AlaskaTempPrecipDischarge.csv")
```


### Create Precipitation only Dataframe with Bins included
```{r, message=FALSE, warning=FALSE}
AlaskaPrecipOnly<-select(AlaskaTempPrecipDischarge, .data$DATE, .data$PRCP, .data$Bin)
```

### Check number of Precipitation NA's and remove them
```{r}
newdata<-AlaskaPrecipOnly
```

### Ensure Date is a date object and not a factor
```{r}
newdata$DATE<-as.Date(newdata$DATE)
```

### Ensure Bin is a factor and not an integer
```{r}
newdata$Bin <- as.factor(newdata$Bin)
```

# Subset the Data by Bin
```{r}
Bin1Precip<-filter(newdata, Bin== 1)
Bin2Precip<-filter(newdata, Bin== 2)
Bin3Precip<-filter(newdata, Bin == 3)
Bin4Precip<-filter(newdata, Bin == 4)
Bin5Precip<-filter(newdata, Bin == 5)
Bin6Precip<-filter(newdata, Bin == 6)
Bin7Precip<-filter(newdata, Bin == 7)
Bin8Precip<-filter(newdata, Bin == 8)
Bin9Precip<-filter(newdata, Bin == 9)
```


### Set Personal Theme
```{r}
gabytheme <- theme_bw(base_size = 22) + 
  theme(plot.title=element_text(face="bold", size="26", color="IndianRed3", hjust=0.5),
        axis.title=element_text(size=22, color="black"),
axis.text = element_text(face="bold", size=16, color = "black"), 
panel.background=element_rect(fill="white", color="darkblue"), 
panel.border = element_rect(color = "black", size = 2),
legend.position = "top", legend.background = element_rect(fill="white", color="black"),
            legend.key = element_rect(fill="transparent", color="NA"))

theme_set(gabytheme)
```



# Plot Bin 1 Precip
```{r, fig.width=12, fig.height=6}
Bin1PrecipPlot<- 
  ggplot(Bin1Precip, aes(x = DATE, y = PRCP)) +
  geom_line() +
  ggtitle("Alaska Precipitation over Time in Latitude Bin 1") +
  labs(x = "", y = expression("Precipitation (in)"))+
  scale_x_date(labels=date_format("%Y"),
               breaks=date_breaks("1 year"))+
  gabytheme
print(Bin1PrecipPlot)

```

# Plot Bin 2 Precip over Time
```{r, fig.width=14, fig.height=6}
Bin2PrecipPlot<- 
  ggplot(Bin2Precip, aes(x = DATE, y = PRCP)) +
  geom_line() +
  ggtitle("Alaska Precipitation over Time in Latitude Bin 2") +
  labs(x = "", y = expression("Precipitation (in)"))+
  scale_x_date(labels=date_format("%y"),
               breaks=date_breaks("1 year"))+
  gabytheme
print(Bin2PrecipPlot)
```
>There is a gap in precipitation data collection for Bin 2 from 1986-12-31 to 1995-10-01

# Plot Bin 3 Precip over Time
```{r, fig.width=30, fig.height=9}
Bin3PrecipPlot<- 
  ggplot(Bin3Precip, aes(x = DATE, y = PRCP)) +
  geom_line() +
  ggtitle("Alaska Precipitation over Time in Latitude Bin 3") +
  labs(x = "", y = expression("Precipitation (in)"))+
  scale_x_date(labels=date_format("%y"),
               breaks=date_breaks("2 years"))+
  gabytheme
print(Bin3PrecipPlot)

```
>Much higher range in precip values-->more rain and snowfall?

# Plot Bin 4 Precip over Time
```{r, fig.width=24, fig.height=6}
Bin4PrecipPlot<- 
  ggplot(Bin4Precip, aes(x = DATE, y = PRCP)) +
  geom_line() +
  ggtitle("Alaska Precipitation over Time in Latitude Bin 4") +
  labs(x = "", y = expression("Precipitation (in)"))+
  scale_x_date(labels=date_format("%y"),
               breaks=date_breaks("1 year"))+
  gabytheme
print(Bin4PrecipPlot)

```

# Plot Bin 5 Precip over Time
```{r, fig.width=24, fig.height=6}
library(scales)
library(ggplot2)
Bin5PrecipPlot<- 
  ggplot(Bin5Precip, aes(x = DATE, y = PRCP)) +
  geom_line() +
  ggtitle("Alaska Precipitation over Time in Latitude Bin 5") +
  labs(x = "", y = expression("Precipitation (in)"))+
  scale_x_date(labels=date_format("%y"),
               breaks=date_breaks("1 year"))+
  gabytheme
print(Bin5PrecipPlot)

```

# Plot Bin 6 Precip over Time
```{r, fig.width=24, fig.height=6}
library(scales)
library(ggplot2)
Bin6PrecipPlot<- 
  ggplot(Bin6Precip, aes(x = DATE, y = PRCP)) +
  geom_line() +
  ggtitle("Alaska Precipitation over Time in Latitude Bin 6") +
  labs(x = "", y = expression("Precipitation (in)"))+
  scale_x_date(labels=date_format("%y"),
               breaks=date_breaks("1 year"))+
  gabytheme
print(Bin6PrecipPlot)

```


# Plot Bin 7 Precip over time
```{r, fig.width=26, fig.height=6}
library(scales)
library(ggplot2)
Bin7PrecipPlot<- 
  ggplot(Bin7Precip, aes(x = DATE, y = PRCP)) +
  geom_line() +
  ggtitle("Alaska Precipitation over Time in Latitude Bin 7") +
  labs(x = "", y = expression("Precipitation (in)"))+
  scale_x_date(labels=date_format("%y"),
               breaks=date_breaks("1 year"))+
  gabytheme
print(Bin7PrecipPlot)
```


# Plot Bin 8 Precip over time
```{r, fig.width=18}
library(scales)
library(ggplot2)
Bin8PrecipPlot<- 
  ggplot(Bin8Precip, aes(x = DATE, y = PRCP)) +
  geom_line() +
  ggtitle("Alaska Precipitation over Time in Latitude Bin 8") +
  labs(x = "", y = expression("Precipitation (in)"))+
  scale_x_date(labels=date_format("%y"),
               breaks=date_breaks("1 year"))+
  gabytheme
print(Bin8PrecipPlot)
```

# Plot Bin 9 Precip over time
```{r, fig.width=18}
library(scales)
library(ggplot2)
Bin9PrecipPlot<- 
  ggplot(Bin9Precip, aes(x = DATE, y = PRCP)) +
  geom_line() +
  ggtitle("Alaska Precipitation over Time in Latitude Bin 9") +
  labs(x = "", y = expression("Precipitation (in)"))+
  scale_x_date(labels=date_format("%y"),
               breaks=date_breaks("1 year"))+
  gabytheme
  
print(Bin9PrecipPlot)
```


# Precip Line Graph but with geom_smooth
```{r, fig.width=13, fig.height=6, warning=FALSE, message=FALSE}
PrecipLinePlot<-ggplot(newdata, aes(x = DATE, y = PRCP, color = Bin))  +
  geom_smooth() +
labs(title="Precipitation Measured over Time", x="Sample Year",
       y="Precipitation (inches)")+ 
  scale_x_date(labels=date_format("%y"),
               breaks=date_breaks("5 years"))+
  gabytheme
print(PrecipLinePlot)
```

## Scatterplot of Full Precipitation Period of Record Colored by Bin
```{r, fig.width=10, fig.height=6}
PrecipScatterplot<-ggplot(newdata, aes(x = DATE, y = PRCP, color = Bin))  +
  geom_point(alpha=0.5, size=2) +
labs(title="Precipitation Measured over Time", x="Sample Year",
       y="Precipitation (inches)")+ 
  scale_x_date(labels=date_format("%y"),
               breaks=date_breaks("5 years"))+
  theme(legend.position="top",
        legend.key.width=unit(1.5, "cm"))+
  gabytheme
print(PrecipScatterplot)
```


## Add Day of Year Column to complete dataframe
```{r, warning=FALSE, message=FALSE}
library(dplyr)
library(lubridate)
newdata <-mutate(newdata, DOY = yday(DATE), Year=year(DATE), Month=month(DATE))
```



# Run a Seasonal Mann-Kendall Test for each Bin

>In many environmental datasets (especially climate and hydrology), we might not expect a linear trend in the response variable over time. In this case, we will need to employ a nonparametric test to determine whether there is a monotonic trend (i.e., consistent increase or decrease but not necessarily linear) over time. We will illustrate a few examples of nonparametric trend tests today with the `trend` package. 
A Mann-Kendall test will analyze whether there is a monotonic trend in the response variable over time. Like a **Mann-Kendall Test**, the **Seasonal Mann-Kendall Test**, or **Hirsch-Slack Test**, analyzes trends in response variables over time. It replaces the traditional Mann-Kendall when there are seasonal trends in a dataset that obscure the overall direction of the trend. It is important to note that "seasonal" does not necessarily equate to actual seasons but can represent any time period within which there are oscillating temporal trends. The test needs at least two distinct seasons to operate.


# Seasonal Mann Kendall Test for Bin 1 Precip
```{r}
# Run a linear interpolation of the dataset to fill in gaps
Bin1Precip.interpolated <- approx(Bin1Precip$DATE,
                                     Bin1Precip$PRCP, 
                                      method = "linear", n = 2492)
# Turn the interpolated dataset into a proper dataframe

Bin1Precip.interpolated  <- do.call(cbind.data.frame, Bin1Precip.interpolated) ##gets us back to data frame
names(Bin1Precip.interpolated ) <- c("DATE", "PRCP")   ###have to rename the columns
Bin1Precip.interpolated$DATE<- as.Date(Bin1Precip.interpolated$DATE, ##have to respecify the Date column as a date. Need to specify the origin too. 
                                            origin = "1970/01/01")


# Convert interpolated dataframe into a time series object for Seasonal Mann-Kendall test

Bin1Precip.timeseries <- ts(Bin1Precip.interpolated$PRCP, 
                                start = c(1967, 11) ,frequency = 12)

### Telling time series to go in a frequency of 12 months per year (monthly), and start in the 11th month of 1967 where the Bin 1 data starts

# Run a Seasonal Mann-Kendall test for Bin 1

Bin1Precip.smktest <- smk.test(Bin1Precip.timeseries)
Bin1Precip.smktest  ###gives us overall trend with a zscore of-.09, p-value of 0.36. So we do not see an increasing trend over time becuse the zscore is negative, but  we have a lot of varability(varsS)

summary(Bin1Precip.smktest)

```
>No seasons here are significant. 


# Seasonal Mann Kendall Test for Bin 2 Precip
```{r}
# Run a linear interpolation of the dataset to fill in gaps
Bin2Precip.interpolated <- approx(Bin2Precip$DATE,
                                     Bin2Precip$PRCP, 
                                      method = "linear", n = 10717)
# Turn the interpolated dataset into a proper dataframe

Bin2Precip.interpolated  <- do.call(cbind.data.frame, Bin2Precip.interpolated) ##gets us back to data frame
names(Bin2Precip.interpolated ) <- c("DATE", "PRCP")   ###have to rename the columns
Bin2Precip.interpolated$DATE<- as.Date(Bin2Precip.interpolated$DATE, ##have to respecify the Date column as a date. Need to specify the origin too. 
                                            origin = "1970/01/01")


# Convert interpolated dataframe into a time series object for Seasonal Mann-Kendall test

Bin2Precip.timeseries <- ts(Bin2Precip.interpolated$PRCP, 
                                start = c(1981, 10) ,frequency = 12)

### Telling it to go in a frequency of 12 months per year (monthly), and start in the 10th month of 1981 where the Bin 2 data starts

## Run a Seasonal Mann-Kendall test for Bin 2

Bin2Precip.smktest <- smk.test(Bin2Precip.timeseries)
Bin2Precip.smktest  ###gives us overall trend with a zscore of-.09, p-value of 0.36. So we do not see an increasing trend over time becuse the zscore is negative, but  we have a lot of varability(varsS)

summary(Bin2Precip.smktest)

```


# Seasonal Mann Kendall Test for Bin 3 Precip
```{r}
# Run a linear interpolation of the dataset to fill in gaps
Bin3Precip.interpolated <- approx(Bin3Precip$DATE,
                                     Bin3Precip$PRCP, 
                                      method = "linear", n = 33461)
# Turn the interpolated dataset into a proper dataframe

Bin3Precip.interpolated  <- do.call(cbind.data.frame, Bin3Precip.interpolated) ##gets us back to data frame
names(Bin3Precip.interpolated ) <- c("DATE", "PRCP")   ###have to rename the columns
Bin3Precip.interpolated$DATE<- as.Date(Bin3Precip.interpolated$DATE, ##have to respecify the Date column as a date. Need to specify the origin too. 
                                            origin = "1970/01/01")


# Convert interpolated dataframe into a time series object for Seasonal Mann-Kendall test

Bin3Precip.timeseries <- ts(Bin3Precip.interpolated$PRCP, 
                                start = c(1915, 6) ,frequency = 12)

### Telling it to go in a frequency of 12 months per year (monthly), and start in the 6th month of 1915 where the Bin 3 data starts

## Run a Seasonal Mann-Kendall test for Bin 3

Bin3Precip.smktest <- smk.test(Bin3Precip.timeseries)
Bin3Precip.smktest  ###gives us overall trend with a zscore of-5.12, p-value of 3.1e-7. So we do not see an increasing trend over time becuse the zscore is negative, but  we have a lot of varability(varsS)

summary(Bin3Precip.smktest)
```



# Seasonal Mann Kendall Test for Bin 4 Precip
```{r}
# Run a linear interpolation of the dataset to fill in gaps
Bin4Precip.interpolated <- approx(Bin4Precip$DATE,
                                     Bin4Precip$PRCP, 
                                      method = "linear", n = 19164)

# Turn the interpolated dataset into a proper dataframe
Bin4Precip.interpolated  <- do.call(cbind.data.frame, Bin4Precip.interpolated) ##gets us back to data frame
names(Bin4Precip.interpolated ) <- c("DATE", "PRCP")   ###have to rename the columns
Bin4Precip.interpolated$DATE<- as.Date(Bin4Precip.interpolated$DATE, ##have to respecify the Date column as a date. Need to specify the origin too. 
                                            origin = "1970/01/01")


# Convert interpolated dataframe into a time series object for Seasonal Mann-Kendall test
Bin4Precip.timeseries <- ts(Bin4Precip.interpolated$PRCP, 
                                start = c(1965, 5) ,frequency = 12)
### Telling it to go in a frequency of 12 months per year (monthly), and start in the 5th month of 1965 where the Bin 4 data starts


## Run a Seasonal Mann-Kendall test for Bin 4

Bin4Precip.smktest <- smk.test(Bin4Precip.timeseries)
Bin4Precip.smktest  ###gives us overall trend with a zscore of 1.793, p-value of 0.07. So we do see an increasing trend over time becuse the zscore is positive, but it is not significant. We do have a lot of varability(varsS)

summary(Bin4Precip.smktest)
```

# Seasonal Mann Kendall Test for Bin 5 Precip
```{r}
# Run a linear interpolation of the dataset to fill in gaps
Bin5Precip.interpolated <- approx(Bin5Precip$DATE,
                                     Bin5Precip$PRCP, 
                                      method = "linear", n = 24312)

# Turn the interpolated dataset into a proper dataframe
Bin5Precip.interpolated  <- do.call(cbind.data.frame, Bin5Precip.interpolated) ##gets us back to data frame

names(Bin5Precip.interpolated ) <- c("DATE", "PRCP")   ###have to rename the columns
Bin5Precip.interpolated$DATE<- as.Date(Bin5Precip.interpolated$DATE, ##have to respecify the Date column as a date. Need to specify the origin too. 
                                            origin = "1970/01/01")


# Convert interpolated dataframe into a time series object for Seasonal Mann-Kendall test
Bin5Precip.timeseries <- ts(Bin5Precip.interpolated$PRCP, 
                                start = c(1952, 4) ,frequency = 12)
### Telling it to go in a frequency of 12 months per year (monthly), and start in the 4th month of 1952 where the Bin 4 data starts


## Run a Seasonal Mann-Kendall test for Bin 5

Bin5Precip.smktest <- smk.test(Bin5Precip.timeseries)
Bin5Precip.smktest  ###gives us overall trend with a zscore of 3.35, p-value of 0.0007. So we do see an significant increasing trend over time becuse the zscore is positive. We do have a lot of varability(varsS)

summary(Bin5Precip.smktest)
```

# Seasonal Mann Kendall Test for Bin 6 Precip
```{r}
# Run a linear interpolation of the dataset to fill in gaps
Bin6Precip.interpolated <- approx(Bin6Precip$DATE,
                                     Bin6Precip$PRCP, 
                                      method = "linear", n = 24583)

# Turn the interpolated dataset into a proper dataframe
Bin6Precip.interpolated  <- do.call(cbind.data.frame, Bin6Precip.interpolated) ##gets us back to data frame

names(Bin6Precip.interpolated ) <- c("DATE", "PRCP")   ###have to rename the columns
Bin6Precip.interpolated$DATE<- as.Date(Bin6Precip.interpolated$DATE, ##have to respecify the Date column as a date. Need to specify the origin too. 
                                            origin = "1970/01/01")


# Convert interpolated dataframe into a time series object for Seasonal Mann-Kendall test
Bin6Precip.timeseries <- ts(Bin6Precip.interpolated$PRCP, 
                                start = c(1951, 6) ,frequency = 12)
### Telling it to go in a frequency of 12 months per year (monthly), and start in the 6th month of 1951 where the Bin 6 data starts


## Run a Seasonal Mann-Kendall test for Bin 6

Bin6Precip.smktest <- smk.test(Bin6Precip.timeseries)
Bin6Precip.smktest  ###gives us overall trend with a zscore of 4.46, p-value of 0.000008. So we do see an significant increasing trend over time becuse the zscore is positive. We do have a lot of varability(varsS)

summary(Bin6Precip.smktest)
```



# Seasonal Mann Kendall Test for Bin 7 Precip
```{r}
# Run a linear interpolation of the dataset to fill in gaps
Bin7Precip.interpolated <- approx(Bin7Precip$DATE,
                                     Bin7Precip$PRCP, 
                                      method = "linear", n = 25586)

# Turn the interpolated dataset into a proper dataframe
Bin7Precip.interpolated  <- do.call(cbind.data.frame, Bin7Precip.interpolated) ##gets us back to data frame

names(Bin7Precip.interpolated ) <- c("DATE", "PRCP")   ###have to rename the columns
Bin7Precip.interpolated$DATE<- as.Date(Bin7Precip.interpolated$DATE, ##have to respecify the Date column as a date. Need to specify the origin too. 
                                            origin = "1970/01/01")


# Convert interpolated dataframe into a time series object for Seasonal Mann-Kendall test
Bin7Precip.timeseries <- ts(Bin7Precip.interpolated$PRCP, 
                                start = c(1948, 5) ,frequency = 12)
### Telling it to go in a frequency of 12 months per year (monthly), and start in the 6th month of 1951 where the Bin 7 data starts


## Run a Seasonal Mann-Kendall test for Bin 7

Bin7Precip.smktest <- smk.test(Bin7Precip.timeseries)
Bin7Precip.smktest  ###gives us overall trend with a zscore of 2.95, p-value of 0.0031. So we do see an significant increasing trend over time becuse the zscore is positive. We do have a lot of varability(varsS)

summary(Bin7Precip.smktest)
```

# Seasonal Mann Kendall Test for Bin 8 Precip
```{r}
# Run a linear interpolation of the dataset to fill in gaps
Bin8Precip.interpolated <- approx(Bin8Precip$DATE,
                                     Bin8Precip$PRCP, 
                                      method = "linear", n = 15169)

# Turn the interpolated dataset into a proper dataframe
Bin8Precip.interpolated  <- do.call(cbind.data.frame, Bin8Precip.interpolated) ##gets us back to data frame

names(Bin8Precip.interpolated ) <- c("DATE", "PRCP")   ###have to rename the columns
Bin8Precip.interpolated$DATE<- as.Date(Bin8Precip.interpolated$DATE, ##have to respecify the Date column as a date. Need to specify the origin too. 
                                            origin = "1970/01/01")


# Convert interpolated dataframe into a time series object for Seasonal Mann-Kendall test
Bin8Precip.timeseries <- ts(Bin8Precip.interpolated$PRCP, 
                                start = c(1976, 10) ,frequency = 12)
### Telling it to go in a frequency of 12 months per year (monthly), and start in the 10th month of 1976 where the Bin 8 data starts


## Run a Seasonal Mann-Kendall test for Bin 8

Bin8Precip.smktest <- smk.test(Bin8Precip.timeseries)
Bin8Precip.smktest  ###gives us overall trend with a zscore of -1.16, p-value of 0.246. So we do see non significant trend over time. We do have a lot of varability(varsS)

summary(Bin8Precip.smktest)
```

# Seasonal Mann Kendall Test for Bin 9 Precip
```{r}
# Run a linear interpolation of the dataset to fill in gaps
Bin9Precip.interpolated <- approx(Bin9Precip$DATE,
                                     Bin9Precip$PRCP, 
                                      method = "linear", n = 13577)

# Turn the interpolated dataset into a proper dataframe
Bin9Precip.interpolated  <- do.call(cbind.data.frame, Bin9Precip.interpolated) ##gets us back to data frame

names(Bin9Precip.interpolated ) <- c("DATE", "PRCP")   ###have to rename the columns
Bin9Precip.interpolated$DATE<- as.Date(Bin9Precip.interpolated$DATE, ##have to respecify the Date column as a date. Need to specify the origin too. 
                                            origin = "1970/01/01")


# Convert interpolated dataframe into a time series object for Seasonal Mann-Kendall test
Bin9Precip.timeseries <- ts(Bin9Precip.interpolated$PRCP, 
                                start = c(1982, 9) ,frequency = 12)
### Telling it to go in a frequency of 12 months per year (monthly), and start in the 9th month of 1982 where the Bin 9 data starts


## Run a Seasonal Mann-Kendall test for Bin 9

Bin9Precip.smktest <- smk.test(Bin9Precip.timeseries)
Bin9Precip.smktest  ###gives us overall trend with a zscore of 10.53, p-value of 2.2e-16. So we do see an increasing significant trend over time. We do have a lot of varability(varsS)

summary(Bin9Precip.smktest)
```

>Interpreting results of the Seasonal Mann-Kendall Test: 

>  + Overall z score and p-value: test the alternative hypothesis that the true change in response variable over time is not equal to zero
  
>  + Monthly z score and p-value: test the alternative hypothesis that the true change in response variable over time for a given month is not equal to zero
  
>  + S: reports trend. A positive value indicates response variable increased over time, and a negative value indicates response variable decreased over time
  





#Determine Water Volume of Precipitation by Bin and Site

# Calculate Cumulative Annual Precipitation for each year by Site and Create Dataframe
```{r}
CumAnnualPrecip<- AlaskaTempPrecipDischarge%>%
  mutate(Year = year(DATE)) %>%
  group_by(site_no, Year)%>%
  summarize(CumPrecip=sum(PRCP, na.rm=TRUE))
         
```

### Load in NWIS_SiteInfo.csv which has Drainage Areas for each Station
```{r}
NWIS_SiteInfo<- read.csv("../DATA/RAW/NWIS_SiteInfo.csv")
```

## Join NWIS_SiteInfo and CumAnnual Precip by Site Number and multiplt cumulative precip for each station and year by drainage area to determine water volume for each year and station
```{r}
PrecipVolume<-CumAnnualPrecip%>%
  merge(NWIS_SiteInfo, by='site_no')%>%
  mutate(Water_Volume=CumPrecip*drain_area_va)%>%
  select(site_no:dec_long_va, drain_area_va, Water_Volume)
```



