---
title: "Temperature Analysis"
author: "Walker Grimshaw"
date: "11/9/2019"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
## install packages and read data
library(tidyverse)
library(lubridate)
library(trend)
library(forecast)
library(tseries)
library(zoo)
Alaska_Data <- read.csv("./DATA/PROCESSED/AlaskaTempPrecipDischarge.csv")
theme_set(theme_bw())
```


```{r}
## OLD CODE NOT NEEDED
## add rolling mean temperature column
Alaska_Data <- mutate(Alaska_Data,
                      TMAX_roll = rollmean(TMAX, 60, fill = NA))

## try plotting monthly averages, all look bad
ggplot(Alaska_Temp_Monthly, aes(x = Month, y = aveTMAX, color = Year)) +
  geom_point(alpha = 0.5) +
  scale_color_viridis_c() +
  facet_wrap(~Bin, nrow = 3, ncol = 3)

```

Though climate change is a complex process, the easiest parameter to estimate the magnitude of climate change is temperature. Figure XXX shows the maximum daily temperature for the nine NOAA sites over each site's period of record. As expected, the range of maximum temperature increases with increasing latitude. Besides site one, all other sites have reasonable overlap in their periods of record and at least cover from 1980 to 2018. There are gaps in the records of multiple sites that require interpolation however. Larger gaps, such as the 1990s gap for site two and the gaps before 1950 for site three, are not filled in this analysis, and instead the earlier time periods are left out of the time series analysis. Shorter data gaps are filled.

```{r}
## Data wrangling
## transform date column from factor to date
Alaska_Data$DATE <- as.Date(Alaska_Data$DATE, format = "%Y-%m-%d")
## Transform Bin column to factor
Alaska_Data$Bin <- as.factor(Alaska_Data$Bin)

## create  NA rows where there are data gaps so lines don't connect
Bin2NA <- c(as.Date("1990-01-01", format = "%Y-%m-%d"), 2, rep(NA, 11))
Bin3NA <- c(as.Date("1945-01-01", format = "%Y-%m-%d"), 3, rep(NA, 11))
Bin3NA_2 <- c(as.Date("1936-01-01", format = "%Y-%m-%d"), 3, rep(NA, 11))
Bin4NA <- c(as.Date("1996-01-01", format = "%Y-%m-%d"), 4, rep(NA, 11))
Bin6NA <- c(as.Date("1995-01-01", format = "%Y-%m-%d"), 6, rep(NA, 11))

## rbind missing date vectors
MissingDates <- as.data.frame(rbind(Bin2NA, Bin3NA, Bin3NA_2, Bin4NA, Bin6NA))
## match names to Alaska Data names
names(MissingDates) <- names(Alaska_Data)
## reformat date column
MissingDates$DATE <- as.Date(MissingDates$DATE)
Alaska_Data_Fig <- rbind(Alaska_Data, MissingDates)
  
## create plot of temperature over time for all 10 latitude bins
ggplot(Alaska_Data_Fig, aes(x = DATE, y = TMAX)) +
  geom_line(alpha = 0.8) +
  facet_wrap(~Bin, nrow = 3, ncol = 3) +
  labs(y = "Maximum Daily Temperature (F)", x = "Date")
```

```{r}
## decomposition of temperature series since the trends are so seasonal
## example for bin 5, must remove NAs
Alaska_Data_5 <- Alaska_Data %>%
  filter(Bin == 5 & is.na(TMAX) == FALSE)
## create time series object of temperature
Alaska_Data_5_ts <- ts(Alaska_Data_5[[9]], frequency = 365)

## decomposition
Alaska_Decomposed <- stl(Alaska_Temp_Monthly_ts, s.window = "periodic")
plot(Alaska5_Decomposed)
```

For each latitude bin, the monthly average maximum temperature was calculated from the daily temperature data. As previously mentioned, the earlier data for sites two and three were discarded and not used for analysis. All other data gaps of at least one year in length were filled using the average maximum temperature for each month over the period of record for each site. However, if there were shorter data gaps of less than one year in length, monthly maximum temperatures were filled using linear interpolation from the nearest neighbor points.

For each latitude bin, a Seasonal Mann-Kendall test was performed on the monthly time series to determine if there had been a change in temperature over the period of record and the directionality of the trend. The Seasonal Mann-Kendall was also used to determine which individual months had statistically significant temperature trends and the directionality of those trends. If there was a statistically significant trend over the period of record, the seasonal sen's slope test was used to determine the average magnitude of the change.

```{r}
## Seasonal Mann-Kendall for whether temperature is changing at each site
## Create monthly dataset for temperature
Alaska_Temp_Monthly <- Alaska_Data %>%
  mutate(Year = year(DATE), Month = month(DATE)) %>%
  group_by(Year, Month, Bin) %>%
  summarize(aveTMAX = mean(TMAX, na.rm = T)) %>%
  mutate(Date = as.yearmon(paste(Year, Month), "%Y %m"))

## Initialize sen's slope vector to be filled in by loop
Temp_Change <- rep(NA, 9)

## Initialize Final Dataframe of monthly data for plotting
Alaska_Temp_Monthly_Model <- data.frame(Date = numeric(),
                                        aveTMAX = numeric(),
                                        Bin = factor())

## Where for loop will start
Site_bin <- 5

for(Site_bin in 1:9){
  
  ## Step 1: filter data by site
  Alaska_Temp_Monthly_Site <- Alaska_Temp_Monthly %>%
    filter(Bin == Site_bin)
  
  ## Use only latest data points when there are large gaps in the data record
  ## identified through examination
  ## would like better way to handle smaller gaps in bins 3, 4, 6, 7
  if(Site_bin == 2){
    Alaska_Temp_Monthly_Site <- Alaska_Temp_Monthly_Site %>%
      filter(Year > 1995 &
               Year < max(Alaska_Temp_Monthly_Site$Year))
  } else if(Site_bin == 3) {
    Alaska_Temp_Monthly_Site <- Alaska_Temp_Monthly_Site %>%
      filter(Year > 1945 &
               Year < max(Alaska_Temp_Monthly_Site$Year))
  } else if(Site_bin == 4) {
    missing_years <- c(1995, 1996)
    Alaska_Temp_Monthly_Site <- Alaska_Temp_Monthly_Site %>%
      filter(Year > min(Alaska_Temp_Monthly_Site$Year)
             & Year < max(Alaska_Temp_Monthly_Site$Year)
             & Year != 1996)
  } else if(Site_bin == 6) {
    missing_years <- c(1995)
    Alaska_Temp_Monthly_Site <- Alaska_Temp_Monthly_Site %>%
      filter(Year > min(Alaska_Temp_Monthly_Site$Year)
             & Year < max(Alaska_Temp_Monthly_Site$Year)
             & Year != 1995)
  } else if(Site_bin == 7) {
    missing_years <- c(2007)
    Alaska_Temp_Monthly_Site <- Alaska_Temp_Monthly_Site %>%
      filter(Year > min(Alaska_Temp_Monthly_Site$Year)
             & Year < max(Alaska_Temp_Monthly_Site$Year)
             & Year != 2007)
  }
  
  ## If missing years, fill in annual gaps. if not, skip to interpolating months
  if(Site_bin %in% c(4,6,7)){
      ## Calculate Mean for each month
    Ave_Temp_Site <- Alaska_Temp_Monthly_Site %>%
      group_by(Month) %>%
      summarize(aveTMAX_month = mean(aveTMAX, na.rm = T)) %>%
    ## add columns to match with entire site data
      transmute(Year = NA, Month = Month, Bin = as.factor(Site_bin),
                aveTMAX = aveTMAX_month)
    
    ## create missing dataframe to fill in annual gaps
    ## need to make the length of the number of missing years
    if(Site_bin == 4){
      Alaska_Temp_Missing <- Ave_Temp_Site %>%
        bind_rows(Ave_Temp_Site)
    } else {
      Alaska_Temp_Missing <- Ave_Temp_Site
    }
    Alaska_Temp_Missing <- Alaska_Temp_Missing %>%
      mutate(Year = rep(missing_years, each = 12),
             Date = as.yearmon(paste(Year, Month), "%Y %m"))
        
    ## fill in gaps and arrange by year and month
    Alaska_Temp_Monthly_Site <- Alaska_Temp_Monthly_Site %>%
      bind_rows(Alaska_Temp_Missing) %>%
      arrange(Year, Month)
  } else if(Site_bin %in% c(1,5,8,9)) {
      Alaska_Temp_Monthly_Site <- Alaska_Temp_Monthly_Site %>%
      filter(Year > min(Alaska_Temp_Monthly_Site$Year)
             & Year < max(Alaska_Temp_Monthly_Site$Year))
  }
  
  ## number of data points for interpolation
  n_months = 12 * (max(Alaska_Temp_Monthly_Site$Year) -
                     min(Alaska_Temp_Monthly_Site$Year) + 1)
  
  Alaska_Temp_Monthly_Interp <- Alaska_Temp_Monthly_Site %>%
    ungroup() %>%
    select(Date, aveTMAX)
  
  ## Interpolate missing months, will return original data frame if complete
  Alaska_Temp_Monthly_Interp <- as.data.frame(approx(Alaska_Temp_Monthly_Interp,
                                                   n = n_months,
                                                   method = "linear")) %>%
    rename(Year = x, aveTMAX = y)
  
  ## create time series object from interpolated temp data
  Alaska_Temp_Monthly_ts <- ts(Alaska_Temp_Monthly_Interp[[2]], frequency = 12)
    
  ## SMK Test for trend
  Alaska_Monthly_Trend <- smk.test(Alaska_Temp_Monthly_ts)
  summary(Alaska_Monthly_Trend)
  ## if significant trend, add slope to slope vector
  ## else add 0 to slope vector
  if (Alaska_Monthly_Trend$p.value > 0.05){
    Temp_Change[Site_bin] <- NA
  } else {
    Slope <- sea.sens.slope(Alaska_Temp_Monthly_ts)
    Temp_Change[Site_bin] <- Slope
  }
  
  ## Add Bin column to interpolated data frame
  Alaska_Temp_Monthly_Interp_Bin <- Alaska_Temp_Monthly_Interp %>%
    mutate(Bin = Site_bin)
  ## rbind Interpolated dataframe to final monthly dataframe for plotting
  Alaska_Temp_Monthly_Model <- rbind(Alaska_Temp_Monthly_Model,
                                     Alaska_Temp_Monthly_Interp_Bin)
}
## coerce Bin column to class factor
Alaska_Temp_Monthly_Model$Bin <- as.factor(Alaska_Temp_Monthly_Model$Bin)

Bin <- seq(1,9)
Sens_Slope_Temp <- as.data.frame(cbind(Bin, Temp_Change))
write.csv(Sens_Slope_Temp, "./Data/PROCESSED/Temperature_Change.csv",
          row.names = F)

```

Figure XXX shows the monthly average maximum temperature for site 5, near Anchorage, over time. The monthly average maximum temperature had a statistically significant trend (Seasonal Mann-Kendall, z=4.87, p<0.001). The blue line shows the magnitude of this trend as calculated by the seasonal sen's slope, indicating a 0.04 degree fahrenheit annual increase in average maximum temperature. It should be noted that the statistical tests used here do not estimate the intercept of the seasonal trend line, so while the figure shows the average seasonal temperature increase, the location of the intercept is only an approximation to assist the visualization.

```{r}
# Make column of predicted temp
Alaska_Temp_Monthly_Site <- Alaska_Temp_Monthly_Site %>%
  mutate(Prediction = 0.0412*(Year - 900))

Alaska_Temp_Monthly_5 <- Alaska_Temp_Monthly_Site
write.csv(Alaska_Temp_Monthly_5, "./DATA/PROCESSED/Alaska_Temp_Monthly_5.csv",
          row.names = F)

ggplot(Alaska_Temp_Monthly_Site, aes(x = Date)) +
  geom_line(aes(y = aveTMAX)) +
  geom_line(aes(y = Prediction), color = "blue", size = 1) +
  labs(y = "Maximum Daily Temperature\nAveraged by Month (F)")
ggplot(Alaska_Temp_Monthly_Interp, aes(x = Year, y = aveTMAX)) +
  geom_line()

```



```{r}
## regression of temperature over bin
Temp_lm <- lm(data = Sens_Slope_Temp, Temp_Change ~ Bin, )
summary(Temp_lm)
```

One possible reason for such little impact of temperature on the date of peak snowmelt is the months in which temperature trends occur. For example, though the overal temperature for Bin 5 shows an increasing trend, the spring months of March, April, and May actually show a cooling trends. Summer and fall months however show increasing temperature trends that outweigh the cooling trends of earlier months. As the sen's slope is the median of all monthly trends, this nuance is not exhibited in the overall temperature trend. Warmer summer and fall seasons may have little impact on the date of peak discharge.

The analyses in this study do not strongly support the hypotheses. The overall temperature trends do not show polar amplification, in part due to a small number of latitude bins and only one temperature site per bin. Similarly, there is no statistically significant relationship between change in discharge over time and latitude. There were only five sites that had both statistically significant trends in temperature and discharge, and these five sites did not show a relationship between temperature and discharge or temperature and day of peak discharge.

Future analsyses should take a more nuanced approach to investigating the impacts of climate change on discharge of Alaskan streams. Temperature data show a need to investigate the seasons that will have direct effects on stream discharge rather than looking at the year as a whole.
