---
output: 
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    number_sections: yes
geometry: margin=2.54cm
title: Climatic Trend Impact on Alaskan Stream Discharge
subtitle: https://github.com/wgrimshaw/Alaska_DBPs.git
author: Gaby Garcia, Walker Grimshaw, Tristen Townsend, Yixin Wen
abstract: "Experimental overview. This section should be no longer than 250 words."
fontsize: 12pt
mainfont: Times New Roman
---

<Information in these brackets are used for annotating the RMarkdown file. They will not appear in the final version of the PDF document>

\newpage
\tableofcontents 
\newpage
\listoftables 
\newpage
\listoffigures 
\newpage

<Setup the global options for the R chunks in your document>

<Note: set up autoreferencing for figures and tables in your document>

```{r setup, include=FALSE, message = FALSE}
knitr::opts_chunk$set(warning = FALSE)
# Set your working directory

# Load your packages
library(tidyverse)
library(lubridate)
library(trend)
library(forecast)
library(tseries)
library(sf)
library(maps)
library(dataRetrieval)
library(viridis)
library(ggplot2)

theme_set(theme_bw())
```


# Research Question and Rationale

<Paragraph detailing the rationale for your analysis. What is the significant application and/or interest in this topic? Connect to environmental topic(s)/challenge(s).>

The climate is changing, in large part due to anthropogenic carbon emmissions. These changes have different magnitudes around the world and local impacts of climate change vary as well. Specifically, climate change is already having greater impacts near the poles than many other parts of the globe. Understanding how climate change is affecting discharge, especially in Alaska, has implications for water management, ecological processes, and the larger global system (if we consider ice-albedo feedback).

## Research Question

To what degree does climate change affect discharge in Alaskan streams and rivers?

<Paragraph detailing your research question(s) and goals. What do you want to find out? Include a sentence (or a few) on the dataset you are using to answer this question - just enough to give your reader an idea of where you are going with the analysis.>

This guiding question encompasses other questions we seek to answer through statistical analyses. If climate change is causing average temperatures to increase, and the magnitude of the increase is greater at higher latitudes, then higher latitude streams should demonstrate greater changes in timing and magnitude of peak flows. This study first seeks to examine if the magnitude of temperature changes increase with increasing latitude in Alaska. By analyzing historical streamflow records, we will investigate whether the magnitude of maximum daily temperature change causes a proportional change in the magnitude and timing of peak streamflow. 

Another aspect of climate change's impacts on streamflow is the potential of melting permafrost or glaciers. This analysis uses cumulative annual streamflow and cumulative annual precipitation to determine if interannual snowpack is melting with increasing average temperatures. If so, we expect the difference between annual precipitation and annual streamflow to increase over time.

\newpage

# Dataset Information

<Information on how the dataset for this analysis were collected, the data contained in the dataset, and any important pieces of information that are relevant to your analyses. This section should contain much of same information as the README file for the dataset but formatted in a way that is more narrative.>

## Discharge

Discharge data were collected from NWIS using the Data Retrieval package. The state of Alaska was divided into 10 bins of equal latitude, and daily discharge data was downloaded for the site in each latitude bin with the greatest number of samples. This dataset includes the site location, daily discharge, and county of the site, among other variables not used in the analysis.

## Temperature and Precipitation

Temperature and precipitation data were downloaded from the National Oceanic and Atmospheric Administration (NOAA) Climate Data Online web portal. As discharge stations do not collect data on temperature and precipitation, the climate data for each latitude bin were downloaded from a station in the same county as each selected discharge station. In each county, daily precipitation, maximum temperature, and minimum temperature, data were downloaded from one station. The criteria for station selection include data extending to the current date, beginning at the earliest date, with at least 80% data coverage. This dataset also includes site location.

<Add a table that summarizes your data structure. This table can be made in markdown text or inserted as a `kable` function in an R chunk. If the latter, do not include the code used to generate your table.>

| Variable | Units (if known) | Type of Variable | Hypothesis | 
|------:|:-----|---------|:------:|
| Discharge | Cubic feet per second | Response | 1a, 1b, 1c |
| Site Number | Latitude/Longitude | Predictor | 1a |
| Date Time | Year/Month/Day | Predictor | 1c |
| Date of First Snowmelt | Year/Month/Day | Predictor | 1a, 1b, 1c |
| Air Temperature | Celsius | Predictor | 1b |
| Precipitation | Millimeters | Predictor | 1a, 1c |
| HUC 8 Watershed Size | Square Meters | Predictor | 1a, 1b | 
| Permafrost Melt | Qualitative | Predictor | 1b | 
| Glacial Coverage/Melting | Qualitative | Predictor | 1b |

\newpage

# Exploratory Data Analysis and Wrangling

<Include R chunks for 5+ lines of summary code (display code and output), 3+ exploratory graphs (display graphs only), and any wrangling you do to your dataset(s).> 

```{r Main Wrangling, message=FALSE, warning=FALSE}
# Import all Data Frames
setwd("/Users/gabrielagarcia/Desktop/Hydrologic Data Analysis/Hydrologic_Data_Analysis/Course_Project/Alaska/DATA/RAW")
NOAATempPrecipBin1<-read.csv("NOAA_Temp_Bin1.csv")
NOAATempPrecipBin2<-read.csv("NOAA_Temp_Bin2.csv")
NOAATempPrecipBin3<-read.csv("NOAA_Temp_Bin3.csv")
NOAATempPrecipBin5<-read.csv("Bin5TempandPrecip.csv")
NOAATempPrecip48910<-read.csv("NOAA_Precip_Temp_Bin_4_8_9_10.csv")

### Add Bin 2 Column to NOAATempPrecipBin2
NOAATempPrecipBin2$Bin<-2

### Remove Tavg Column from Precip 
library(dplyr)
NOAATempPrecip48910<-select(NOAATempPrecip48910,-c(8))

### Add Bin Columns with values for 4, 8, and 9
library(plyr)
NOAATempPrecip48910$Bin<-revalue(NOAATempPrecip48910$STATION,
                              c('USW00026510'='4', 'USC00500464'='8','USW00027502'='9'))

### Upload Bins 6 and 7 
setwd("/Users/gabrielagarcia/Desktop/Hydrologic Data Analysis/Hydrologic_Data_Analysis/Course_Project/Alaska/DATA/PROCESSED")
NOAATempPrecipBin67<-read.csv("NOAA_Temp_Precip_Bin_6_7.csv")

### Remove Snow Column from Precip Temp 6 & 7
NOAATempPrecipBin67<-select(NOAATempPrecipBin67,-c(8))

# Fix Date Issues in NOAA Bin 1


##Tell R to change factor to a date, in the format Month/day/year
NOAATempPrecipBin1$DATE <- as.Date(NOAATempPrecipBin1$DATE, format = "%m/%d/%y") 

# Changes the format of the "datetime" column to a two digit year, a two digit month, and the day as a number
NOAATempPrecipBin1$DATE <- format(NOAATempPrecipBin1$DATE, "%y%m%d")

 
create.early.dates <- (function(d) {
       paste0(ifelse(d > 181231,"19","20"),d)
       })

### This will input the datetime column into the function we created above for each row of the column
NOAATempPrecipBin1$DATE<- create.early.dates(NOAATempPrecipBin1$DATE)

### Changes the date time column into a different format to a 4 digit year, a two digit month, and the day as
NOAATempPrecipBin1$DATE <- as.Date(NOAATempPrecipBin1$DATE, format = "%Y%m%d")

# Fix Date Issues in NOAA Bin 5

## Tell R to change factor to a date, in the format Month/day/year
NOAATempPrecipBin5$DATE <- as.Date(NOAATempPrecipBin5$DATE, format = "%m/%d/%y") 

## Changes the format of the "datetime" column to a two digit year, a two digit month, and the day as a number
NOAATempPrecipBin5$DATE<- format(NOAATempPrecipBin5$DATE, "%y%m%d")

## The 181231 is the last date of 2018. If d is over this value, a 19 is returned, If D<the number, a 20 is returned.
create.early.dates <- (function(d) {
       paste0(ifelse(d > 181231,"19","20"),d)
       })

### This will input the datetime column into the function we created above for each row of the column
NOAATempPrecipBin5$DATE<- create.early.dates(NOAATempPrecipBin5$DATE)

### Changes the date time column into a different format to a 4 digit year, a two digit month, and the day as
NOAATempPrecipBin5$DATE <- as.Date(NOAATempPrecipBin5$DATE, format = "%Y%m%d") 

# RBind all TempPrecip Data
TempPrecip<-rbind(NOAATempPrecipBin1, NOAATempPrecipBin2, NOAATempPrecipBin3, NOAATempPrecipBin5, NOAATempPrecipBin67, NOAATempPrecip48910)

# Import USGS Discharge Data
setwd("/Users/gabrielagarcia/Desktop/Hydrologic Data Analysis/Hydrologic_Data_Analysis/Course_Project/Alaska/DATA/RAW")
NWIS_Discharge<-read.csv("NWIS_Discharge.csv")

### Convert site_no to a factor
NWIS_Discharge$site_no<-as.factor(NWIS_Discharge$site_no)

## Add Bins to Discharge Data
library(plyr)
NWIS_Discharge$Bin<-revalue(NWIS_Discharge$site_no,
                              c('15297680'='1', '15297610'='2','15072000'='3', '15052500'='4', '15276000'='5', '15304000'='6', '15514000'='7', '15453500'='8', '15908000'='9', '15896000'='10'))

### Rename "Date" to "DATE" in NWIS_Discharge dataframe so both columns match
names(NWIS_Discharge)[3]<-"DATE"

# Join TempPrecip and NWIS_Discharge Data Frames by Date and Bin 
AlaskaTempPrecipDischarge <- TempPrecip %>%
  merge(NWIS_Discharge, by = c('DATE', 'Bin')) %>%
  rename(Discharge = X_00060_00003) %>%
  select(DATE:Discharge)
```




The first step of the analysis process is determining the presence and magnitude of temperature change across the latitude range of Alaska. The daily maximum temperature data show no obvious trends over time for any of the sites. Furthermore, the sites differ substantially in the period of record of their temperature data. Sites with no recent observations may need to be dropped from analysis, unless nearby streams have similar periods of record.

```{r Temperature Graph, echo=FALSE, fig.cap= "Daily maximum temperature for the period of record of each latitude bin. There are no obvious trends, and the period of record differs substantially for each site."}

AlaskaTempPrecipDischarge <- read.csv("./DATA/PROCESSED/AlaskaTempPrecipDischarge.csv")
## transform date column from factor to date
AlaskaTempPrecipDischarge$DATE <- as.Date(AlaskaTempPrecipDischarge$DATE,
                                          format = "%Y-%m-%d")
## Transform Bin column to factor
AlaskaTempPrecipDischarge$Bin <- as.factor(AlaskaTempPrecipDischarge$Bin)

## create plot of temperature over time for all 10 latitude bins
ggplot(AlaskaTempPrecipDischarge, aes(x = DATE, y = TMAX)) +
  geom_line(alpha = 0.8) +
  facet_wrap(~Bin, nrow = 3, ncol = 3) +
  labs(y = "Maximum Daily Temperature (F)", x = "Date")
```

# Full Precipitation Period of Record over Time, divided by Bin 
```{r, fig.width=10, fig.height=6, echo = FALSE}
library(scales)

### Load Data
AlaskaTempPrecipDischarge <- read.csv("./DATA/PROCESSED/AlaskaTempPrecipDischarge.csv")
AlaskaPrecipOnly<-select(AlaskaTempPrecipDischarge, .data$DATE, .data$PRCP, .data$Bin)
newdata<-na.omit(AlaskaPrecipOnly)

### Create and set theme
gabytheme <- theme_bw(base_size = 22) + 
  theme(plot.title=element_text(face="bold", size="26", color="IndianRed3", hjust=0.5),
        axis.title=element_text(size=22, color="black"),
axis.text = element_text(face="bold", size=16, color = "black"), 
panel.background=element_rect(fill="white", color="darkblue"), 
panel.border = element_rect(color = "black", size = 2),
legend.position = "top", legend.background = element_rect(fill="white", color="black"),
            legend.key = element_rect(fill="transparent", color="NA"))

theme_set(gabytheme)


### Plot Precipitation 
PrecipBoxplot<-ggplot(newdata, aes(x = DATE, y = PRCP, color = Bin))  +
  geom_point(alpha=0.5, size=2) +
labs(title="Full Precipitation Period of Record Measured over Time by Bin", x="Sample Year",
       y="Precipitation (inches)")+ 
  scale_x_date(labels=date_format("%y"),
            breaks=date_breaks("5 years"))+
  scale_fill_manual(guide=guide_legend(keyheight=unit(4, units="mm")))+
  #gabytheme
print(PrecipBoxplot)
```


<<<<<<< HEAD
Give a general image of how discharge changes over time in different latitudes.
```{r Discharge, plot discharge over time for 10 bins}
=======
```{r Discharge, plot discharge over time for 10 bins, echo=FALSE}
>>>>>>> 6361a6ebcd717e909f77edcf77a429d25d408d69
#import the data
AlaskaTempPrecipDischarge <- read.csv("./DATA/PROCESSED/AlaskaTempPrecipDischarge.csv")
# transform bin into factor
AlaskaTempPrecipDischarge$Bin <- as.factor(AlaskaTempPrecipDischarge$Bin)
#Extract date and discharge data
AlaskaDischarge <- AlaskaTempPrecipDischarge%>%
  select("DATE", "Bin","STATION","NAME", "site_no", "Discharge")
#drop na
AlaskaDischarge%>%drop_na()
# plot discharge over time for 10 latitude bins
discharge_plot <- ggplot(AlaskaDischarge,aes(x = DATE, y = Discharge))+
  geom_line()+
  facet_wrap(~Bin, nrow = 3, ncol = 3)+
  labs(x = "Date", y =expression("Discharge (ft"^3*"/s)"))
print(discharge_plot)
```


Discharge is to be used as a proxy for snowmelt. The data was loaded and wrangled to include the year, month, day, day of year, and week in order to better analyze if and how the day of first snowmelt is changing. The data was filtered for weeks 1-30 of each year (mid-winter to mid-summer), to minimize the amount of data being plotted in exploratory graphs. The mean discharge for each day of year through time by latitude bin was plotted. This served to provide insight into roughly when the day of first snowmelt was for each site, and to see how different sites varied. The data was then filtered to look at a single latitude bin, in this case bin 6, and see how discharge changed by the day of year across the period of record. In order to better visual any trend, bin 6 was filtered for data on a ten year frequency for the past fifty years (1979, 1989, 1999, 2009, 2019). After looking at the various exploratory graphs, it was decided to proceed with using peak discharge as a proxy for peak snowmelt, rather than try to determine when the first snowmelt of each year was occuring. 


```{r Snowmelt Change, echo = FALSE}

#Load the data
AlaskaTempPrecipDischarge <- read.csv("./DATA/PROCESSED/AlaskaTempPrecipDischarge.csv")
#Convert DATE column to class 'Date'
AlaskaTempPrecipDischarge$DATE <- ymd(AlaskaTempPrecipDischarge$DATE)

#Add date columns
Snowmelt.Discharge <- AlaskaTempPrecipDischarge %>% mutate(YEAR = year(DATE), 
                        MONTH = month(DATE), 
                        DAY = day(DATE), WEEK = week(DATE), 
                        DOY = yday(DATE))

Snowmelt.Discharge.Spring <- Snowmelt.Discharge %>% filter(WEEK < 30)

#Plot mean discharge for each DOY through time by bin
Snowmelt.Discharge.Mean <- Snowmelt.Discharge.Spring %>% group_by(Bin, DOY) %>%
  mutate(Dis.Mean = mean(Discharge))

dis.doy.mean <- ggplot(Snowmelt.Discharge.Mean) +
  geom_line(aes(x = DOY, y = Dis.Mean, color = as.factor(Bin))) +
  labs(x = "Day of Year", y = "Mean Discharge, cfs", color = "Latitude Bin") +
  scale_color_viridis_d() +
  theme(legend.position = "top")
print(dis.doy.mean)


#Make dataframes for bin analysis
#Filter for only needed columns and bin
Snowmelt.Discharge.Bin6 <- Snowmelt.Discharge.Spring %>% select(DATE, YEAR, MONTH, DAY, WEEK, DOY, Bin, NAME, LATITUDE, LONGITUDE, Discharge) %>%
  filter(Bin == 6)

#Latitude Bin 6 - discrete years - every 10
Snowmelt.Discharge.Bin6.10Year <- Snowmelt.Discharge.Bin6 %>% 
  filter(YEAR == 1979 | YEAR == 1989 | YEAR == 1999 | YEAR == 2009 | YEAR == 2019)

dis.doy.6.10 <- ggplot(Snowmelt.Discharge.Bin6.10Year) +
  geom_line(aes(x = DOY, y = Discharge, color = as.factor((YEAR)))) +
  labs(x = "Day of Year", y = "Discharge, cfs", color = "Year") +
  scale_color_viridis_d() +
  theme(legend.position = "top")
print(dis.doy.6.10)

```


<Include text sections to accompany these R chunks to explain the reasoning behind your workflow, and the rationale for your approach.>


\newpage

# Analysis
<Include R chunks for 3+ statistical tests (display code and output) and 3+ final visualization graphs (display graphs only).>

## Long-term Temperature and Discharge Trends

A seasonal Mann-Kendall test was used to determine if there was a long-term trend in maximum temperature and stream discharge for each latitude.

A seasonal Sen's Slope quantifies the magnitude of the temperature and discharge trends.

<<<<<<< HEAD

From general discharge image of 10 bins, Bin 6 seems to have seasonal change. Thus apply smk.test on bin 6 discharge data to see if there's a seasonal trend. adf.test is also applied to see if the data is stationary. If it is, use auto.arima function to find the best fit model to predict the future discharge and plot the graph. 
```{r Discharge}
=======
```{r Discharge, echo=FALSE}
>>>>>>> 6361a6ebcd717e909f77edcf77a429d25d408d69
# Time series for Bin 6
AlaskaDischarge_6 <- AlaskaDischarge %>% 
  filter(Bin == 6)%>%
  drop_na()

AlaskaDischarge_6.Monthly <- AlaskaDischarge_6%>%
  mutate(Year = year(DATE), Month = month(DATE))%>%
  group_by(Year,Month)%>%
  summarise(Discharge = mean(Discharge))

AlaskaDischarge_6_ts.Monthly <- ts(AlaskaDischarge_6.Monthly[[3]], frequency = 12)

AlaskaDischarge_6_trend.Monthly <-smk.test(AlaskaDischarge_6_ts.Monthly) 
AlaskaDischarge_6_trend.Monthly
summary(AlaskaDischarge_6_trend.Monthly)
# p-value = 0.0001581 < 0.05, which means bin 6 has obvious seasonal change

# test if the data is stationary
adf.test(AlaskaDischarge_6_ts.Monthly, alternative = "stationary")
# since the p-value is 0.01 < 0.05, the data is stationary

#run the arima function and search for best fit
auto.arima(AlaskaDischarge_6_ts.Monthly, trace = TRUE)

# create an object that defines the best fit model
fit <- arima(AlaskaDischarge_6_ts.Monthly, c(1, 0, 0),seasonal = list(order = c(1, 1, 0), period = 12)) 

# make a prediction into the future
AlaskaDischarge_6_prediction <- predict(fit, n.ahead = 10*12)

# plot future predictions
ts.plot(AlaskaDischarge_6_ts.Monthly, AlaskaDischarge_6_prediction$pred, lty = c(1, 3))

```

## Predictability of snowmelt

Is there a change over time of the period when snow melts? Peak discharge will be used as a proxy for peak snowmelt. At each latitude bin, the day of year of peak snowmelt will be found for each year in the period of record. If thare are any gaps in the data, it will be addressed through linear interpolation. Once the day of year of peak snowmelt is found, a dataframe of 'Year' and 'Day of Year of Peak Snowmelt' will be created, and a time series will be run to see if there is any significant change at each site in the day of peak snowmelt. 


<Include text sections to accompany these R chunks to explain the reasoning behind your workflow, rationale for your approach, and the justification of meeting or failing to meet assumptions of tests.>


\newpage

# Summary and Conclusions
<Summarize your major findings from your analyses. What conclusions do you draw from your findings? Make sure to apply this to a broader application for the research question you have answered.>



